---
title: "Modelando con R"
author: "David Suárez <david.suarez@yahoo.com>"
date: "28 de abril de 2018"
output: revealjs::revealjs_presentation
---

# Modelación

En general hay dos pasos en todo proceso de modelación:

1.  Escoges una familia de modelos `y = a_1 * x + a_2` o `y = a_1 * x ^ a_2`.

1.  Generas un modelo ajustado, como `y = 3 * x + 7` o `y = 9 * x ^ 2`, donde `a_1` y `a_2` toman ciertos valores.

A través de este proceso obtendremos el mejor modelo, pero no necesariamente el verdadero. Como dice un aforismo de George Box:

> All models are wrong, but some are useful.

La cita completa es:

> Now it would be very remarkable if any system existing in the real world 
> could be exactly represented by any simple model. However, cunningly chosen 
> parsimonious models often do provide remarkably useful approximations. For 
> example, the law PV = RT relating pressure P, volume V and temperature T of 
> an "ideal" gas via a constant R is not exactly true for any real gas, but it 
> frequently provides a useful approximation and furthermore its structure is 
> informative since it springs from a physical view of the behavior of gas 
> molecules.
> 
> For such a model there is no need to ask the question "Is the model true?". 
> If "truth" is to be the "whole truth" the answer must be "No". The only 
> question of interest is "Is the model illuminating and useful?".

La meta de la modelación no es descubrir la verdad, sino obtener una buena aproximación que sea útil. 

### Prerequisitos: modelr

```{r setup, message = FALSE}
#library(tidyverse)

library(modelr)
options(na.action = na.warn)
```

## Un modelo simple

modelr::sim1

```{r}
ggplot(sim1, aes(x, y)) + 
  geom_point()
```

Pareciera haber un patrón en nuestros datos, y pareciera ser lineal: `y = a_0 + a_1 * x`.

```{r}
models <- tibble(
  a1 = runif(250, -20, 40),
  a2 = runif(250, -5, 5)
)

ggplot(sim1, aes(x, y)) + 
  geom_abline(aes(intercept = a1, slope = a2), data = models, alpha = 1/4) +
  geom_point() 
```

Pero, qué tan bueno es cada uno de esos modelos:

```{r, echo = FALSE}
dist1 <- sim1 %>% 
  mutate(
    dodge = rep(c(-1, 0, 1) / 20, 10),
    x1 = x + dodge,
    pred = 7 + x1 * 1.5
  )

ggplot(dist1, aes(x1, y)) + 
  geom_abline(intercept = 7, slope = 1.5, colour = "grey40") +
  geom_point(colour = "grey40") +
  geom_linerange(aes(ymin = y, ymax = pred), colour = "#3366FF") 
```

Para calcular esas funciones, requerimos convertir nuestra familia de modelos en una función:

```{r}
model1 <- function(a, data) {
  a[1] + data$x * a[2]
}
model1(c(7, 1.5), sim1)
```
